HMMotion: Using tracking data to predict coverage strategies in American Football

### 2024-12-29 submission.Rmd
.
sub

@@

## Introduction
Deciding between man and zone coverage is one of the most critical strategic choices a defensive coordinator must make before each offensive play in American football. While experienced offensive coordinators and quarterbacks often rely on visual cues to identify these defensive schemes, the increasing availability of player tracking data offers a new avenue to uncover and analyze these tactics. A notable example is Amazon's NFL Next Gen Stats model, which delivers predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns. These predictions, however, are often limited to analyzing specific plays without incorporating pre-snap motion which is a crucial element of modern offensive strategies.

@@

<center>
  <img src="prime_coverage_prediction.jpg" width="600" height="400" alt="Coverage Prediction">
</center>

@@

##### Feature engineering
To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data. In particular, we conducted the following feature engineering steps: ROBERT

>>>>

```{r colorize,echo=FALSE}
#| echo: false
# colorize text: use inline as `r colorize(text, color)` to print `text` in a given `color`
# can also be used to color a color name, as in r colorize("red")`
colorize <- function(text, color) {
  if (missing(color)) color <- text
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, text)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, text)
  } else text
}
```

@@

## Introduction
Deciding between man and zone coverage is one of the most critical strategic choices a defensive coordinator must make before each offensive play in American football. While experienced offensive coordinators and quarterbacks often rely on visual cues to identify these defensive schemes, the increasing availability of player tracking data offers a new avenue to uncover and analyze these tactics. A notable example is Amazon's NFL Next Gen Stats model, which delivers predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns. These predictions, however, are often limited to analyzing specific plays without incorporating pre-snap motion which is a crucial element of modern offensive strategies. `r colorize("Ich weiß nicht, ob ich das so formulieren würde... Wir wissen nicht was deren Modell alles inkludiert. Es kann gut sein, dass sie auch tracking und daher auch premotion Daten nehmen. Oder glaubst Du ist das eine Expertise von einer Person mit domain knowledge und kein statistisches Modell?","red")`

@@

<center>
  <img src="prime_coverage_prediction.jpg" width="500" height="250" alt="Coverage Prediction">
</center>

@@

##### Feature engineering
To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data. In particular, we conducted the following feature engineering steps: We first consider all 11 players on each side of the field and compute features related to the convex hull of the positions of the players. In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e. the width of the hull) and the largest $x$ distance (i.e. the length of the hull). In addition, we select the 5 most relevant players on each side of the field (for offense: we omit the offensive line and the QB; for defense: `r colorize("TODO, Rouven","red")`) and use their standardized $x$ and $y$ coordinates as covariates. (`r colorize("Distances, orientation?? Appenix??","red")`)
